---
title: 'Project 2: Modeling and Evaluation'
subtitle: '<p>CSE6242 - Data and Visual Analytics</p><p>Due: Friday, April 21, 2017
  at 11:59 PM UTC-12:00 on T-Square</p>'
output:
  pdf_document: default
  html_document: default
---

# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, as well as a PDF export of it (as **pr2.pdf**) which should include all of that plus output, plots and written responses for each task.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r}
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
library(ggplot2)
library(GGally)
library(stringr)
library(reshape)
library("tokenizers")
library(discretization)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: 
GGally
tokenizers
reshape
discretization

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# TODO: Remove all rows from df that do not correspond to movies
df_movie=df[df$Type=="movie",]
nrow(df_movie)
```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r, echo=FALSE, warning=FALSE}
# TODO: Remove rows with missing Gross value
df_movie_gross=df_movie[!is.na(df_movie$Gross),]
nrow(df_movie_gross)
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r, message=FALSE, warning=FALSE}
# TODO: Exclude movies released prior to 2000
df_movie_2000=df_movie_gross[df_movie_gross$Year>=2000,]
nrow(df_movie_2000)
```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r, message=FALSE, warning=FALSE}
# TODO: Remove mismatched rows
test=df_movie_2000[setdiff(rownames(df_movie_2000), rownames(df_movie_2000[is.na(df_movie_2000$Released),])),]
test_match=test[abs((test$Year-as.numeric(substr(test$Released, 1, 4))))<=1,]
test_dismatch=df_movie_2000[setdiff(rownames(df_movie_2000), rownames(df_movie_2000[rownames(test_match),])),]
df_movie_match=test_match
nrow(df_movie_match)
```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r, message=FALSE, warning=FALSE}
# TODO: Exclude the `Domestic_Gross` column
df_movie_match <- df_movie_match[,!(colnames(df_movie_match) %in% c("Domestic_Gross"))] 
```

## 6. Process `Runtime` column

```{r}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes
# Check the text feature of the Runtime column
unique(str_extract_all(df_movie_match$Runtime, "[A-z]+"))
```
```{r}
#Replace df$Runtime with a numeric column containing the runtime in minutes
for(i in c(1:nrow(df_movie_match))){
  if(is.na(str_extract_all(df_movie_match[i,"Runtime"], "[A-z]+"))){
    df_movie_match[i,"Runtime"]=as.numeric(str_extract_all(df_movie_match[i,"Runtime"], "[[:digit:]]+"))
  }else if(length(str_extract_all(df_movie_match[i,"Runtime"], "[A-z]+")[[1]])==2){
    df_movie_match[i,"Runtime"]=(as.numeric(str_extract_all(df_movie_match[i,"Runtime"], "[[:digit:]]+")[[1]][1]))*60 +(as.numeric(str_extract_all(df_movie_match[i,"Runtime"], "[[:digit:]]+")[[1]][2]))
  }else if(str_extract_all(df_movie_match[i,"Runtime"],"[A-z]+")=="h"){
    df_movie_match[i,"Runtime"]=(as.numeric(str_extract_all(df_movie_match[i,"Runtime"], "[[:digit:]]+")))*60
  }else{
    df_movie_match[i,"Runtime"]=as.numeric(str_extract_all(df_movie_match[i,"Runtime"], "[[:digit:]]+"))
  }
}

```

```{r}
df_movie_match$Runtime=as.numeric(df_movie_match[,"Runtime"])
```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

```{r}
# TODO(optional): Additional preprocessing
```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
cat("Dataset has", dim(df_movie_match)[1], "rows and", dim(df_movie_match)[2], "columns", end="\n", file="")
colnames(df_movie_match)
```

# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, use the following evaluation procedure every time:

1. Randomly divide the rows into two sets of sizes 5% and 95%.
2. Use the first set for training and the second for testing.
3. Compute the Root Mean Squared Error (RMSE) on the train and test sets.
4. Repeat the above data partition and model training and evaluation 10 times and average the RMSE results so the results stabilize.
5. Repeat the above steps for different proportions of train and test sizes: 10%-90%, 15%-85%, ..., 95%-5% (total 19 splits including the initial 5%-95%).
6. Generate a graph of the averaged train and test RMSE as a function of the train set size (%).

You can define a helper function that applies this procedure to a given model and reuse it.

# Tasks

Each of the following tasks is worth 20 points. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

```{r}
#Build a function to split data based on percentage criteria
sample_and_shuffle=function(data, sample_perc){

smp_size <- floor(sample_perc * nrow(data))
train_ind <- sample(seq_len(nrow(data)), size = smp_size)

train <- data[train_ind, ]
test <- data[-train_ind, ]

new_data=list(train, test)
  return(new_data)
}
```

```{r}
#Build a function to compute RMSE
rmse=function(lmmodel){
  rss=c(crossprod(lmmodel$residuals))
  mse=rss/length(lmmodel$residuals)
  result_rmse=sqrt(mse)
  return(result_rmse)
}
```

```{r}
#Build a function to compute RMSE for test data
test_rmse=function(resid)
{
  result_rmse=(mean( (resid)^2,  na.rm =TRUE ))^0.5 
  return(result_rmse) 
}

```

```{r}
# Build a function to generate columns string for modeling
model_criteria=function(col_list){
columns="Gross"
i=1
for(column in col_list)
{
  if(column!="Gross"){
  if(i==1)
  {
  columns=paste (columns,column,sep = "~", collapse = NULL)}
  else{
  columns=paste (columns,column,sep = "+", collapse = NULL)  
  }
  
  i=i+1}
}
return(columns)
}
```




```{r}
#Build a function to train based on training data and test data split criteria
train_by_split=function(split, col_list,dataset){
data=sample_and_shuffle(dataset,split)
training_data=data[[1]]
test_data=data[[2]]
fit=lm(eval(parse(text=model_criteria(col_list))), training_data)
residual=test_data$Gross-predict(fit, newdata=test_data)
training_rmse=rmse(fit)
testing_rmse=test_rmse(residual)
return(c(training_rmse,testing_rmse))
}
```


```{r}
#Build a function to compute average RMSE
train_avg=function(split,col_list,dataset,repeat_times=10){
  result= data.frame(matrix(ncol = 2, nrow = 0))
  colnames(result)=c("RMSE_train", "RMSE_test")
  for(i in c(1:repeat_times)){
    
    result=rbind(result, as.data.frame(matrix(train_by_split(split,column_list,dataset), ncol = 2, nrow = 1)))
    
  }
  colnames(result)=c("RMSE_train", "RMSE_test")
  
  return(colMeans(result))
}
```



```{r}
#Build a function to compute RMSE over different training sizes
train_over_size=function(size_step, column_list,dataset){
  size=0.05
  result= data.frame(matrix(ncol = 3, nrow = 0))
  colnames(result)=c("Train_Perc", "RMSE_train", "RMSE_test")
  while(size<1){
  result=rbind(result,as.data.frame(matrix(unlist(list(size,train_avg(size,column_list,dataset))),1,3)) ) 
  size=size+size_step
  }
    colnames(result)=c("Train_Perc", "RMSE_train", "RMSE_test")
  return(result)
}
```


## 1. Numeric variables

Use linear regression to predict `Gross` based on all available _numeric_ variables.

```{r, message=FALSE, warning=FALSE}
# TODO: Build & evaluate model 1 (numeric variables only)
# 1.Select the numeric columns (In this case, integer are also considered numeric)
column_list=colnames(df_movie_match[,sapply(df_movie_match, class) %in% c("numeric","integer")])
print(column_list)
```

```{r, message=FALSE, warning=FALSE}
# Train over different sizes and get the result
train_over_size(0.05,column_list,df_movie_match)
```


```{r, message=FALSE, warning=FALSE}
#Plot the result
data <- melt(train_over_size(0.05,column_list,df_movie_match), id.vars = "Train_Perc")
ggplot(data, aes(x = Train_Perc, y = value, colour = variable)) + geom_line()
```



**Q**: List all the numeric variables you used.

**A**: "Year" ,"Runtime" ,"imdbRating" ,"imdbVotes" ,"tomatoMeter", "tomatoRating", "tomatoReviews", "tomatoFresh","tomatoRotten", "tomatoUserMeter", "tomatoUserRating","tomatoUserReviews", "Budget", "Gross", "Date"  

## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).

```{r}
# TODO: Build & evaluate model 2 (transformed numeric variables only)
# 1. Check the data distribution of the numeric values
summary(df_movie_match[,colnames(df_movie_match[,sapply(df_movie_match, class) %in% c("numeric","integer")])])
```

```{r}
# 2. Logarithmize and discretize some numeric columns
# make a new data
df_movie_match_new=df_movie_match
df_movie_match_new$tomatoUserReviews= cut_number(df_movie_match[,"tomatoUserReviews"], n = 6) 
df_movie_match_new$tomatoRotten=cut_number(df_movie_match[,"tomatoRotten"], n = 6) 
df_movie_match_new$tomatoFresh=cut_number(df_movie_match[,"tomatoFresh"], n = 6) 
df_movie_match_new$tomatoReviews=cut_number(df_movie_match[,"tomatoReviews"], n = 6)
df_movie_match_new$imdbVotes=log(df_movie_match[,"imdbVotes"])

```

```{r}
# 3. Check again
summary(df_movie_match_new[,colnames(df_movie_match[,sapply(df_movie_match, class) %in% c("numeric","integer")])])
```

```{r, message=FALSE, warning=FALSE}
#Plot the result
result=train_over_size(0.05,column_list,df_movie_match_new)
print(result)
data <- melt(result, id.vars = "Train_Perc")
ggplot(data, aes(x = Train_Perc, y = value, colour = variable)) + geom_line()
```

**Q**: Explain which transformations you used and why you chose them.

**A**: I categorized columns:
       1. tomatoUserReviews
       2. tomatoRotten
       3. tomatoFresh
       4. tomatoReviews
       and logarithmized column:
       5. imdbVotes
       
       Because reviews can be categorize into different categories and the imdbVotes is very skewed, Logarithmization can make it normal.
       

## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)

# Convert Awards to 2 numeric columns: wins and nominations
# Process the "Awards" column and fill in the "wins" and "nominations" columns
award_convert=function(df){
for (i in c(1:nrow(df))){
    win=0
    nomination=0
  if(grepl("Won",df[i, "Awards"])){
    win=as.numeric(str_extract(str_extract(df[i,"Awards"], "Won ([\\d]+)"), "[[:digit:]]+"))
    win=win+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) win"), "[[:digit:]]+"))
    nomination=nomination+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) nomination"), "[[:digit:]]+"))
  }else if(grepl("Nominated",df[i, "Awards"])){
    nomination=as.numeric(str_extract(str_extract(df[i,"Awards"], "Nominated for ([\\d]+)"), "[[:digit:]]+"))
    win=win+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) win"), "[[:digit:]]+"))
    nomination=nomination+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) nomination"), "[[:digit:]]+"))
  }else{
    win=win+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) win"), "[[:digit:]]+"))
    nomination=nomination+as.numeric(str_extract(str_extract(df[i,"Awards"], "([\\d]+) nomination"), "[[:digit:]]+"))
  }
    df[i,"wins"]=win
    df[i, "nominations"]=nomination
    if(is.na(df[i,"wins"])){df[i,"wins"]=0}
    if(is.na(df[i,"nominations"])){df[i,"nominations"]=0}
}
return(df)
}


# Build a function to replace Genre with a collection of binary columns
# We list all of the genres
genre_vector=function(df){
unique(unlist(tokenize_regex(df$Genre,pattern=", ")))
#Combine the genre with the original dataset
column_genre_names=unique(unlist(tokenize_regex(df$Genre,pattern=", ")))
genre=data.frame(matrix(nrow=nrow(df), ncol=length(column_genre_names)))
colnames(genre)=column_genre_names
genre[]=0
for(val in column_genre_names){
  genre[grep(val, df[, "Genre"]), val]=1
}
genre_vectored=cbind(df, genre)
return(genre_vectored)
}
```


```{r}
column_list_new=colnames(df_movie_match[,sapply(df_movie_match, class) %in% c("character","Date")])
print(column_list_new)
```

```{r}
df_movie_match_new=df_movie_match
df_movie_match_new=award_convert(df_movie_match_new)
df_movie_match_new=genre_vector(df_movie_match_new)
```
```{r}
colnames(df_movie_match_new)
```

```{r}
df_movie_match_new_temp=df_movie_match_new[,c(37,39:64)]
```

```{r}
sapply(df_movie_match_new_temp, class)
```

```{r}
column_list=colnames(df_movie_match_new_temp[,sapply(df_movie_match_new_temp, class) %in% c("numeric","integer")])
df_movie_match_new_temp=df_movie_match_new_temp[,sapply(df_movie_match_new_temp, class) %in% c("numeric","integer")]
column_list[5]="SciFi"
colnames(df_movie_match_new_temp)=column_list
column_list=colnames(df_movie_match_new_temp[,sapply(df_movie_match_new_temp, class) %in% c("numeric","integer")])
column_list=column_list[-26]
print(column_list)
```


```{r, message=FALSE, warning=FALSE}
#Plot the result
result=train_over_size(0.05,column_list,df_movie_match_new_temp)
print(result)
data <- melt(result, id.vars = "Train_Perc")
ggplot(data, aes(x = Train_Perc, y = value, colour = variable)) + geom_line()
```

**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: I choose the categrical variable:
       1. Awards
       2. Genre
       Because it contains many different types and can be converted to binary columns.
       The column `Genre` represents a list of genres associated with the movie in a string format. Write code to parse each text string into a binary vector with 1s representing the presence of a genre and 0s the absence, and add it to the dataframe as additional columns. Then remove the original `Genre` column.

For example, if there are a total of 3 genres: Drama, Comedy, and Action, a movie that is both Action and Comedy should be represented by a binary vector <0, 1, 1>. Note that you need to first compile a dictionary of all possible genres and then figure out which movie has which genres (you can use the R `tm` package to create the dictionary).
       
       The variable `Awards` describes nominations and awards in text format. Convert it to 2 numeric columns, the first capturing the number of wins, and the second capturing nominations. Replace the `Awards` column with these new columns, and then study the relationship of `Gross` revenue with respect to them.

Note that the format of the `Awards` column is not standard; you may have to use regular expressions to find the relevant values. Try your best to process them, and you may leave the ones that don't have enough information as NAs or set them to 0s.
       
       

## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r, message=FALSE, warning=FALSE}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)

df_movie_match_model4=cbind(df_movie_match[,sapply(df_movie_match, class) %in% c("numeric","integer")], df_movie_match_new_temp[,c(2:25)])
column_list=colnames(df_movie_match_model4)
print(column_list)

```
```{r, message=FALSE, warning=FALSE}
#Plot the result
result=train_over_size(0.05,column_list,df_movie_match_model4)
print(result)
data <- melt(result, id.vars = "Train_Perc")
ggplot(data, aes(x = Train_Perc, y = value, colour = variable)) + geom_line()

```


## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r, message=FALSE, warning=FALSE}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)
# In the dataset, column "Comedy" indicates if the genre is Comedy or not. it counts as an additional feature
#Plot the result
result=train_over_size(0.05,column_list,df_movie_match_model4)
print(result)
data <- melt(result, id.vars = "Train_Perc")
ggplot(data, aes(x = Train_Perc, y = value, colour = variable)) + geom_line()

```
```

**Q**: Explain what new features you designed and why you chose them.

**A**: I choose if the genre is comedy. Since it's an interaction feature.
